Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-04-30 writing/embedded_meta_learning.tex
Clarity, clarity, clarity

\begin{document}
\title{Embedded meta-learning}
\author{Andrew K. Lampinen}
\date{}
\maketitle

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is the separation of knowledge within deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between these. That is, neural networks knowledge about what is being computed is inaccessible to those computations. By contrast, humans are frequently consciously aware of both our representations of input, the things we can do with them, and the relationships between data and tasks. This is true for domains from game playing to natural language understanding. \par
One of the abilities that this grants humans is rapid adaptation of our behavior. The problem of rapid learning has been partially addressed by meta-learning systems. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a functional perspective on meta-learning and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language just as data can.}. By treating tasks and data as functions, we can conceptually think of both data \emph{and} learned behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits such as a useful framework for continual learning. \par
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relateively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par

@@

\end{tikzpicture}
\caption{Schematic of our general architecure. Blocks with solid edges denote deep networks with learnable parameters, dashed edges represent inputs, outputs, embeddings, etc., and $f$ is a deep network with parameters specified by $\mathcal{H}$.} \label{architecture_fig}
\end{figure}
More formally, we take a \textbf{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. \par
We can then interpret most machine learning tasks as a mapping of functions to functions. These functions could represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, or they could be functions that operate on functions themselves. Under this perspective, learning tasks and learning to flexibly map between tasks, or learning to map from language to tasks, are all the same type of problem. \par
Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Targets are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. (Targets do not necessarily need to be output-like, e.g. in our RL tasks, we use (action, outcome) tuples as ``targets.'') \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to the (embedded) targets. These transformations are performed by a system with the following components:  
\begin{itemize}
\item $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which takes a set of (input embedding, target embedding) pairs and produces a function embedding. 
\item $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which takes a function embedding and produces a set of parameters. 
\item $f: Z \rightarrow Z$ -- the transformation, implemented by a deep network with parameters specified by $\mathcal{H}$.
\end{itemize}
See Fig. \ref{architecture_fig} for a schematic of the architecture. \par

>>>>

\begin{document}
\title{Embedded Meta-Learning: Toward more flexible deep-learning models}
\author{Andrew K. Lampinen}
\date{}
\maketitle

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrasts, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep{?????}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a functional perspective on meta-learning and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating tasks and data as functions, we can conceptually think of both data \emph{and} learned behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits such as a useful framework for continual learning. \par
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par

@@

\end{tikzpicture}
\caption{Schematic of our general architecure. Blocks with solid edges denote deep networks with learnable parameters, dashed edges represent inputs, outputs, embeddings, etc., and $f$ is a deep network with parameters specified by $\mathcal{H}$.} \label{architecture_fig}
\end{figure}
More formally, we take a \emph{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. \par
We can then interpret most machine learning tasks as a mapping of functions to functions. These functions could represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, or they could be functions that operate on functions themselves. Under this perspective, learning tasks and learning to flexibly map between tasks, or learning to map from language to tasks, are all the same type of problem. \par
Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Targets are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. (Targets do not necessarily need to be output-like, e.g. in our RL tasks, we use (action, outcome) tuples as ``targets.'') \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to the (embedded) targets. These transformations are performed by a system with the following components:  
\begin{itemize}
\item $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which takes a set of (input embedding, target embedding) pairs and produces a function embedding. 
\item $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which takes a function embedding and produces a set of parameters. 
\item $f: Z \rightarrow Z$ -- the transformation, implemented by a deep network with parameters specified by $\mathcal{H}$.
\end{itemize}
See Fig. \ref{architecture_fig} for a schematic of the architecture. \par
